Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-07-16 Paper.Rmd
ceder-creek description fig and description added

\begin{figure}[!h]
\textbf{ \hspace{4.3 cm} (A) \hspace{4.3 cm} (B) \hspace{4 cm} (C) } \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=5 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
\includegraphics[width=5 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}\\
\end{center}
\textbf{ \hspace{4.3 cm} (D) \hspace{4.3 cm} (E) \hspace{4 cm} (F) } \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=5 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
\includegraphics[width=5 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs. ln(mass) of contrasts, \textbf{(C)} contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits.
Lower panel \textbf{(D - F)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for last two datasets (see \nameref{Model_selection} of manuscript). \label{fig_biv_raw_cop}}
\end{center}
\end{figure}

@@

For all $3$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis) 
in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
highest AIC-weight based asymmetric copulas - which answers the first part of Q1. 
Each dataset presented there, has significant upper tail dependence 
compared to Normal as indicated by negative AIC-weights model averaged 
(lower-upper) tail dependence and very low ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics, which answers the second and third part of Q1
(For detailed results of each dataset see Tables \ref{SM-tab_soilCN_asym} - \ref{SM-tab_contrasts_67_asym} and 
Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_contrasts_67_nonparam}).

@@

tab_biivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
knitr::kable(tab_bivar_summary, 
             format='pandoc',
             caption = "Summary of results for bivariate data-sets.\\label{tab_bivar_summary}",
             booktabs=TRUE)
```

>>>>

\begin{figure}[!h]
\textbf{ \hspace{3.2 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{3.2 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs. ln(mass) of contrasts, \textbf{(C)} contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits. \textbf{(D)} above-ground plant biomass data vs. Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
Lower panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript). \label{fig_biv_raw_cop}}
\end{center}
\end{figure}

@@

For all $4$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis, Ceder-creek) 
in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
highest AIC-weight based asymmetric copulas - which answers the first part of Q1. 
Among four datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal. It was indicated by negative AIC-weights model averaged 
(lower-upper) tail dependence and very low ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics for first three datasets and positive AIC-weights model averaged 
(lower-upper) tail dependence and very high ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of Q1
(For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}).

@@

tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
knitr::kable(tab_bivar_summary, 
             format='pandoc',
             caption = "Summary of results for bivariate data-sets.\\label{tab_bivar_summary}",
             booktabs=TRUE)
```

### 2018-07-16 SupportingInformation.Rmd
ceder-creek description fig and description added

@@

  if(ploton==T){
    if(rawplot==T){
      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col=rgb(0,0,1,0.3),pch=19,cex.lab=1.5)
      mtext(paste0("Year=",year))
    }else{
      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col=rgb(1,0,0,0.3),pch=19,cex.lab=1.5)
      mtext(paste0("Year=",year))
    }
  }

@@

  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
}
par(op)
dev.off()
#----------------------------------------------------------------------------------------
cop_HB_all_yr<-vector("list",length(yr))
names(cop_HB_all_yr)<-yr

@@

  s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
  cop_HB_all_yr[[i]]<-s$cop_tab_dbp[,2:3]
}
par(op)
dev.off()

@@

@@

@@

@@

}
#----------------------------------
#calling the above function on 3 multivariable datasets
m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
m2<-bivar_tab_summary_fn(r=BivMS_BMR,rs=stat_BMR_ga)
m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67,rs=stat_contrasts_67_ga)
tab_bivar_summary<-cbind(m1,m2,m3)
colnames(tab_bivar_summary)<-c("SoilCN","BMR","Anolis")
saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
```

>>>>

These biodiversity experiment data sets, for [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and for [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120), are collected by Tilman et al. over many years. We used thsese datasets [@Tilman2001; @Tilman2006] to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m),
for year 2000.

@@

  if(ploton==T){
    if(rawplot==T){
      op<-par(mar=c(5.1, 4.1, 1.5, 2.1))
      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col="red",pch=1,cex.lab=1.6,cex.axis=1.5)
      mtext(paste0("Year=",year),cex=1.5)
      par(op)
    }else{
      op<-par(mar=c(5.1, 4.1, 1.5, 2.1))
      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",col="red",pch=1,cex.lab=2,cex.axis=1.5,
         xlim=c(0,1),ylim=c(0,1),asp=1)
      rect(0,0,1,1)
      lines(c(0,1),c(0,1),type='l')
      mtext(paste0("Year=",year),cex=1.5)
      par(op)
    }
  }

@@

  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
}
par(op)
dev.off()
#-----------------------
#raw data plot for year2000
pdf("./Results/Ceder_creek_results/raw_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
dev.off()
#----------------------------------------------------------------------------------------
cop_HB_all_yr<-vector("list",length(yr))
names(cop_HB_all_yr)<-yr

@@

  s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
  cop_HB_all_yr[[i]]<-s$cop_tab_dbp[,2:3]
}
par(op)
dev.off()
#-------------------------
#cop data plot for year2000
pdf("./Results/Ceder_creek_results/cop_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
cc_cop_2000<-s$cop_tab_dbp[,2:3]
dev.off()
#------------------

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\vspace{-0.2 cm}
\begin{center}
\includegraphics[width=3 in]{./Results/Ceder_creek_results/raw_HB_2000.pdf}
\includegraphics[width=3 in]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
\caption[\textbf{(A)} Plot of the plant-aboveground avg. Biomass vs. avg. Shannon's index (H). \textbf{(B)} Copula plot of plant-aboveground avg. Biomass vs. avg. Shannon's index.]{\textbf{(A)} Plot of the plant-aboveground avg. Biomass vs. avg. Shannon's index (H). \textbf{(B)} Copula plot of plant-aboveground avg. Biomass vs. avg. Shannon's index. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(cc_cop_2000)`$ is the number of (9m by 9m)plots. (see \nameref{MT-Model_selection} of manuscript). \label{fig_raw_cop_cc2000}}
\end{center}
\end{figure}

@@

<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **Ceder-creek dataset** for year 2000. \\label{tab_cc2000fit}",
             booktabs=T)
```

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for Ceder-creek data with $numbin=10$ for year 2000]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{Ceder-creek data}. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).  When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than. \label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

```{r tab_cc_asym_2000, echo=F, results="markup"}
tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
knitr::kable(tab_cc_asym_2000,
             format="pandoc",
             caption = "**Ceder-creek data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_cc_asym_2000}",
             booktabs=T)
```

@@

}
#----------------------------------
#calling the above function on 3 multivariable datasets
m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
m2<-bivar_tab_summary_fn(r=BivMS_BMR,rs=stat_BMR_ga)
m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67,rs=stat_contrasts_67_ga)
m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
tab_bivar_summary<-cbind(m1,m2,m3,m4)
colnames(tab_bivar_summary)<-c("SoilCN","BMR","Anolis","Ceder-creek")
saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
```
